---
title: "UAS PDS"
author: "Muhammad Jamie Rofie Quality"
date: "2022-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Intro

## Calling the Libraries

```{r}
library(factoextra)
library(cluster)
library(ppclust)
library(clusterSim)
library(REdaS)
library(fclust)
library(ggplot2)
library(dplyr)
library(broom)
library(ggdendro)
library(readxl)
library(fmsb)
library(NbClust)
library(RColorBrewer)
library(gridExtra)
library(robustHD)
library(clValid)
library(dendextend)
library(psych)
```

## Importing the Data

```{r}
data <- read_excel("Statistik Kriminal Weighted.xlsx")
View(data)
data <- as.data.frame(data)
rownames(data) <- data$Polda
colnames(data) <- c('Daerah', 'murder', 'assault', 'rape', 'abduction', 'theftv', 'theftwv', 'drugs', 'fnc', 'capo')  ## renaming the variables
data <- data[-35,-1] ## dropping row 1 (Provinsi)
head(data)
```


```{r}
plot(data)
```

```{r}
boxplot(data) ## data needn't to be scaled as the variables are measured in the similar way
data <- as.data.frame(data)
```

```{r}
par(mfrow=c(3,3))
hist(data[,1], main = "Murder")
hist(data[,2], main = "Assault")
hist(data[,3], main = "Rape")
hist(data[,4], main = "Abduction")
hist(data[,5], main = "Theft With Violence")
hist(data[,6], main = "Theft Without Violence")
hist(data[,7], main = "Drugs Crime")
hist(data[,8], main = "Fraud and Corruption Crime")
hist(data[,9], main = "Crime Against Public Orders")
```

# Preparing the Data

## KMO Test to Determine the Adequacy of the Data
```{r}
KMOS(data)
```
## Bartlett Test of Spherecity to Determine the Multicollinearity
```{r}
bart_spher(data)
```

## Principal Components Analysis
```{r}
pca1 <- princomp(data, scores=TRUE, cor=TRUE)
summary(pca1)
```
### Scree Plot of Eigenvalue
```{r}
eig <- get_eigenvalue(pca1)
eig
fviz_eig(pca1)
```


### Loadings
```{r}
loadings(pca1)
data.pca <- pca1$scores [,1:3]
head(data.pca)
```


## Internal validation
```{r}
intern <- clValid(data.pca, nClust = 2:6, clMethods = c("hierarchical", "kemans","fanny"), validation ="internal")
summary(intern)
```

## Stability Validation
```{r}
stabil <- clValid(data.pca, nClust = 2:6, clMethods = c("hierarchical","kmeans"), validation="stability")
optimalScores(stabil)
```

## Elbow Method with WSS
```{r}
par(mfrow=c(2,2))

# Elbow method
fviz_nbclust(data.pca, kmeans, method = "wss") +
  labs(subtitle = "Elbow method")
```

## Silhouette method
```{r}
par(mfrow=c(2,2))

# Elbow method
fviz_nbclust(data.pca, kmeans, method = "silhouette") +
    geom_vline(xintercept = 2, linetype = 3)+
  labs(subtitle = "Silhouette method")
```

# K-Means Clustering
## Clustering Process
```{r}
set.seed(123)
km_out = kmeans(data.pca, 2)
fviz_cluster(km_out, data = data.pca)
```

## Attaching the K-Means Result to the Dataset
```{r}
res <- cbind(data.pca, km_out$cluster)
head(res)
```
## Cluster Characteristics
```{r}
new <- aggregate(res[,-ncol(res)], list(res[,ncol(res)]), mean)
new
# Set graphic colors
library(RColorBrewer)
coul <- brewer.pal(2, "Set2")
colors_border <- coul
library(scales)

# If you remove the 2 first lines, the function compute the max and min of each variable with the available data:
radarchart( new[,-1], axistype=0 , maxmin=F,
    #custom polygon
    pcol=colors_border, pfcol=coul, plwd=4, plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="black", cglwd=0.8, 
    #custom labels
    vlcex=0.8 
    )

# Add a legend
legend(x=0.7, y=1, legend = new$Group.1, bty = "n", pch=20, col=coul, text.col = "grey", cex=1.2, pt.cex=3)
```
```{r}
library(RColorBrewer)
coul <- brewer.pal(2, "Set2")
colors_border <- coul
library(scales)
fviz_cluster(km_out, data = data.pca)
```

# Hierarchical Clustering

## Clustering Process with Four Linkage Method

```{r}
hc_complete = hclust(dist(data.pca), method = "complete")
hc_average = hclust(dist(data.pca), method = "average")
hc_single = hclust(dist(data.pca), method = "single")
hc_ward = hclust(dist(data.pca), method = "ward.D")
```

### Agglomerative Coefficient
```{r}
hc1 <- agnes(data.pca, method ="complete")
hc2 <- agnes(data.pca, method ="single")
hc3 <- agnes(data.pca, method ="average")
hc4 <- agnes(data.pca, method ="ward")

hc1$ac
hc2$ac
hc3$ac
hc4$ac
```

## Dendogram of Each Method
```{r}
plot_complete = ggdendrogram(hc_complete, rotate = FALSE, size = 2) + labs(title = "Complete Linkage") 
plot_average = ggdendrogram(hc_average, rotate = FALSE, size = 2) + labs(title = "Average Linkage")
plot_single = ggdendrogram(hc_single, rotate = FALSE, size = 2) + labs(title = "Single Linkage")
plot_ward = ggdendrogram(hc_ward, rotate = FALSE, size = 2) + labs(title = "Ward Linkage")

grid.arrange(plot_complete)
grid.arrange(plot_average)
grid.arrange(plot_single)
grid.arrange(plot_ward)
```

### Clustering Result of Single Linkage Method
```{r}
plot(hc_single, cex=0.6)
rect.hclust(hc_single, k=2, border = 2:5)
```

### Clustering Result of Average Linkage Method
```{r}
plot(hc_average, cex=0.6)
rect.hclust(hc_average, k=2, border = 2:5)
```

### Clustering Result of Complete Linkage Method
```{r}
plot(hc_complete, cex=0.6)
rect.hclust(hc_complete, k=2, border = 2:5)
```

### Clustering Result of Ward Method
```{r}
plot(hc_ward, cex=0.6)
rect.hclust(hc_ward, k=2, border = 2:5)
```

## Attaching Ward's Result to the Dataset
```{r}
clust <- cutree(hc_ward, 2)
clust
```

```{r}
hc_res <- cbind(data.pca, clust)
hc_res
```

## Cluster Characteristics with Ward's Hierarchical Clustering method
```{r}
new <- aggregate(hc_res[,-ncol(hc_res)], list(hc_res[,ncol(hc_res)]), mean)
new
# Set graphic colors
library(RColorBrewer)
coul <- brewer.pal(2, "Set2")
colors_border <- coul
library(scales)

# If you remove the 2 first lines, the function compute the max and min of each variable with the available data:
radarchart( new[,-1], axistype=0 , maxmin=F,
    #custom polygon
    pcol=colors_border, pfcol=coul, plwd=4, plty=1,
    #custom the grid
    cglcol="grey", cglty=1, axislabcol="black", cglwd=0.8, 
    #custom labels
    vlcex=0.8 
    )

# Add a legend
legend(x=0.7, y=1, legend = new$Group.1, bty = "n", pch=20, col=coul, text.col = "grey", cex=1.2, pt.cex=3)
```

#Fuzzy C-Means Clustering

##Clustering Process

```{r}
res.fcm <- fcm(data.pca, centers=2)
as.data.frame(res.fcm$u)[1:34,]
```
## Cluster Characteristic with Fuzzy C-Means Clustering
```{r}
res.fcm$v0
```

```{r}
res.fcm
```
## Attaching The Result to the Dataset
```{r}
data.fcm <- cbind(data.pca, res.fcm$cluster)
data.fcm
```

```{r}
plotcluster(res.fcm, cp=1, trans=TRUE)
res.fcm2 <- ppclust2(res.fcm, "fanny")
fviz_cluster(res.fcm2, data = data.pca, 
  ellipse.type = "convex",
  palette = "jco",
  repel = TRUE)
```

# Cluster Characteristics with Criminality Variables

```{r}
clust1 <- data[-c(25,29,31,33,34),]
summary(clust1)
```
```{r}
clust2 <- data[c(25,29,31,33,34),]
summary(clust2)
```
